---
title: "Version_1.12_diff_simulations"
author: "Yi Zheng"
date: "12/16/2019"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls(all.names = TRUE))
library(mvtnorm)
library(numDeriv)
library(Matrix)
library(ggplot2)
library(splines)
library(tidyverse)
library(gganimate)
library(kableExtra)
```

```{r read pep data, message=FALSE, warning=FALSE}
PEP<-read_csv("ev_prefiltering.csv")
PEP<-PEP$PEP
anyNA(PEP)
```

```{r parameters}
n<-100
num_of_data <- n
p <- 50
num_it <- 25
degf <- 6

```


# Simulation set up

Here I use $\sqrt{(1-PEP)}$ to denoise the simulation data.


```{r simulation set up}

#set.seed(123)
set.seed(100)
temp <- sample(20:30,p, replace = TRUE)
power <- runif(p,2,4)

pep_sim<-matrix(sample(PEP,n*p,replace = TRUE), nrow = n, ncol = p)
#pep_sim<-matrix(0, nrow = n, ncol = p)
#s <- runif(num_of_data,0, 1)
s <- rbeta(num_of_data, 8, 2)
x<-matrix(nrow = n, ncol = p)
y<-matrix(nrow = n, ncol = p)
z<-matrix(nrow = n, ncol = p)
y1<-matrix(nrow = n, ncol = p)
x1<-matrix(nrow = n, ncol = p)

#x is the reference RT
for (i in 1:p) {
  x[,i] <- 3*(temp[i]-10)^2*s^power[i]/100 + 10
}

prior <- array(dim = c(n,p,n))
for (i in 1:n) {
  for (j in 1:p) {
    prob=rep(pep_sim[i,j]/(n-1),n)
    prob[i]=1-pep_sim[i,j]
    prior[i,j,]=prob
    prob1 <- rep((1-sqrt(prob[i]))/(n-1),n)
    prob1[i] <- sqrt(prob[i])
    class=rmultinom(1, 1, prob1)
    y[i,j]=t(class)%*%x[,j]+rnorm(1,mean = 0,sd=0.1)
    z[i,j]=which(class==1)
  }
}

# for (i in 1:(p-1)) {
#   for (j in (i+1):p) {
#     plot<-ggplot(data=as.data.frame(y))+geom_point(aes(x=y[,i],y=y[,j]))
#     print(plot)
#   }
# }

for (i in 1:p) {
  y1[,i] <- 3*(temp[i]-10)^2*sort(s)^power[i]/100 + 10-mean(y[,i])
  x1[,i] <- 3*(temp[i]-10)^2*s^power[i]/100 + 10-mean(y[,1])
}


```


# Alignment Initialization

```{r alignment 1 initialization}
y <- scale(y, scale = F)

svdx <- svd(y)
d<-svdx$v

# ggplot(data=as.data.frame(y))+geom_point(aes(x=y[,1],y=y[,2]))+geom_abline(intercept = 0, slope = d[2,1]/d[1,1])
# ggplot(data=as.data.frame(y))+geom_point(aes(x=y[,1],y=y[,3]))+geom_abline(intercept = 0, slope = d[3,1]/d[1,1])
# ggplot(data=as.data.frame(y))+geom_point(aes(x=y[,2],y=y[,3]))+geom_abline(intercept = 0, slope = d[3,1]/d[2,1])

## plot projections of each point onto line
proj1 <- y%*%d[,1]%*%t(d[,1])


ak_initial<-function(fi,mu,d){
  ak<-t(d)%*%(fi-mu)/(t(d)%*%d)
  return(ak)
}
n=nrow(y)
p=ncol(y)
vk<-rep(1/n,n)
mu<-colMeans(y)
Fun<-proj1#n by p matrix representing fj(ak)
a<-apply(Fun, 1, ak_initial, mu=mu,d=d[,1])

a <- (a-min(a))/(max(a)-min(a))

aini<-a
alist<-list()

#sigma<-t(replicate(n,diag(cov(y))))#should be n by p (col for each ak)
sigma<-matrix(0.25^2,nrow = n,ncol = p)



```


# Alignment Iteration

```{r alignment 2 iteration}
Q<-function(a,fits,W,X){#f is a list of fits, is empty for first iteration
  n=length(a)
  p<-ncol(X)
  Fun<-matrix(data = NA,nrow = n,ncol = p)
  if(length(fits)==0){#first iteration
    # mu<-colMeans(X)
    # d<-svdx$v[,1]
    # Fun<-t(d%*%t(a)+mu)
    # Sigvec<-diag(cov(X))
    # Qval=0
    # for (k in 1:n) {
    #   S=(t(t(X)-Fun[k,]))^2
    #   second_term<-rep(0.5*sum(log(Sigvec)),n)
    #   Qval<-Qval+sum(S%*%(-1/Sigvec)-second_term)
    # }
  }else{
    for (j in 1:p) {
      Fun[,j]=predict(fits[[j]],a)$y
    }
    Qval<-0
    for (k in 1:n) {
      S=W[,,k]*(t(t(X)-Fun[k,]))^2
      #S=(t(t(X)-Fun[k,]))^2
      Sigvec=rep(0.25^2,p)#p by 1 vector of sigma_j^2(ak)
      second_term<-sum(W[,,k]%*%log(Sigvec)/2)
      Qval<-Qval+sum(S%*%(-1/Sigvec))/2-second_term
    }
  }
  return(-Qval)
}


#step2 Iteration
#likelihood function
plot_dat <- cbind(Fun,rep(0,n),y)
u=0
fits<-list()
SMtx<-list()
#PenMtx<-list()
# Qval<-rep(0,num_it+1)
# Qval[1]<--Q(a,fits,W=NA,x)#Return value of Q is the negative of Q value in EM
# Qval1<-Qval[1]
for (h in 1:num_it) {
  # Qval0<-Qval1
  W2<-array(dim = c(n,p,n))#modify the calculation
  for (i in 1:n) {
    for (j in 1:p) {
      gyij<-0
      for (k1 in 1:n) {
        gyij<-gyij+dnorm(y[i,j],Fun[k1,j],sd = 0.25)*prior[i,j,k1]
      }
      for (k in 1:n) {
        loggyijak<-log(dnorm(y[i,j],mean = Fun[k,j],sd = 0.25))
        #gy(y_h)
        W2[i,j,k]<-loggyijak-log(gyij)+log(prior[i,j,k])
      }
    }
  }
  W2=exp(W2)
  if(h!=1){
    a<-nlm(f=Q, p=a, fits=fits, W=W2, X=y,iterlim = 1)$estimate
  }
  #now update f and sigma and vk
   a2<-max(a)
   a1<-min(a)
  weights<-t(apply(W2, 2:3, sum))
  
  for (j in 1:p) {
    Dmtx<-diag(1/weights[,j])#it's the inverse of D
    ybar<-t(W2[,j,])%*%y[,j]
    yvalue<-Dmtx%*%ybar
    fit <- smooth.spline(x=a, y=yvalue, w=weights[,j], df=degf)##fixed df because as iteration goes on the fit will be interplotation.
    Fun[,j]=predict(fit,a)$y
    fits[[j]]=fit
  }
  
################################################################################  
  for (k in 1:n) {
    S=(t(t(y)-Fun[k,]))^2
    #sigma[k,]=colSums(W2[,,k]*S)/colSums(W2[,,k])#p by 1 vector of sigma_j^2(ak)
  }
  vk<-weights/(n*p)
  u<-u+1
  # plot data and the principal curve for a sequence of a's
  # plot(y[,1], y[,2], bty='n',
  #      xlab=expression(x[1]),
  #      ylab=expression(x[2]))
  # legend("topleft", legend=c("Proj 1"), bty="n")
  # points(x=Fun[,1],y=Fun[,2],col="red")
  # 
  # points(x=y[2,1],y=y[2,2],col="blue")
  
  plot_dat <- rbind(plot_dat,cbind(Fun,rep(h,n),y))
   a<-(a-a1)/(a2-a1)
   alist[[h]]<-a
}

```

# Iteration Process In animation

```{r plots of iteration process}
plot_dat <- as.data.frame(plot_dat)
xlim <- c(min(plot_dat$V1-1), max(plot_dat$V1)+1)
ylim <- c(min(plot_dat$V2-1), max(plot_dat$V2)+1)
plt <- ggplot(data = plot_dat, aes(x=V1, y=V2))+geom_point(color = "red")+scale_x_continuous(limits = xlim)+scale_y_continuous(limits = ylim)+geom_point(aes(x=V52,y=V53))
#+geom_path(data = data.frame(x=y1[,1],y=y1[,2],aes(x=x,y=y)))
plt + transition_states(states = V51, state_length = 0.2, transition_length = 0.2)+ggtitle('Now showing {closest_state}',
          subtitle = 'Frame {frame} of {nframes}')
```

# S value and a value comparison

```{r analysis of results: a values and s values}
hist(a)
hist(s)


pt.dat <- data.frame(y1=y[,1],y2=y[,2],y3=y[,3],a=a,s=s,f1=Fun[,1],f2=Fun[,2],f3=Fun[,3])

ggplot(data = pt.dat)+geom_point(aes(x=s,y=y1))+geom_point(aes(x=s,y=f1,colour="red"))
ggplot(data = pt.dat)+geom_point(aes(x=s,y=y2))+geom_point(aes(x=s,y=f2,colour="red"))
ggplot(data = pt.dat)+geom_point(aes(x=s,y=y3))+geom_point(aes(x=s,y=f3,colour="red"))

ggplot(data = pt.dat)+geom_point(aes(x=a,y=y1))+geom_point(aes(x=a,y=f1,colour="red"))
ggplot(data = pt.dat)+geom_point(aes(x=a,y=y2))+geom_point(aes(x=a,y=f2,colour="red"))
ggplot(data = pt.dat)+geom_point(aes(x=a,y=y3))+geom_point(aes(x=a,y=f3,colour="red"))
```

# The effect of the model


```{r analysis: comparison}
updated_true <- matrix(0, nrow = nrow(y), ncol = ncol(y))
mtx_updated <- matrix(0, nrow = nrow(y), ncol = ncol(y))
updated_id <- matrix(0, nrow = nrow(y), ncol = ncol(y))
for (j in 1:p) {
  for (i in 1:n) {
    updated_true[i,j] <- W2[i,j,z[i,j]]
    mtx_updated[i,j] <- max(W2[i,j,])
    updated_id[i,j] <- if(max(W2[i,j,])>0.99) which.max(W2[i,j,]) else i
    #updated_id[i,j] <- which.max(W2[i,j,])
  }
}

mtx_updated_true_pep <- 1-updated_true
mtx_updated_pep <- 1-mtx_updated

sum(as.numeric(pep_sim<0.01))
sum(as.numeric(mtx_updated_true_pep<0.01))
sum(as.numeric(mtx_updated_pep<0.01))
```

# Explanatory plots

```{r analysis of results: explanatory plots}
#new_pep <- as.numeric(mtx_updated_true_pep)
new_pep <- as.numeric(mtx_updated_pep)
old_pep <- as.numeric(pep_sim)
#plot1<-ggplot(data = data.frame(new=new_pep,old=old_pep))+geom_point(aes(x=new,y=old))
#plot1+geom_abline(slope = 1,intercept = 0)

#plot1+xlim(0,0.05)+geom_abline(slope = 1,intercept = 0)

id_tf<-as.numeric(z==matrix(rep(1:n,p),nrow = n))

pep_dat<-data.frame(new=new_pep,old=old_pep,id_tf=id_tf, pair=rep(1:n,p))

plot2<-ggplot(data = pep_dat,aes(x=new,y=old))+geom_point(aes(colour=id_tf))
plot2+geom_abline(slope = 1,intercept = 0)
plot2+xlim(0,0.05)+geom_abline(slope = 1,intercept = 0)


#plot2+geom_abline(slope = 1,intercept = 0)+geom_line(aes(group=pair))

#filtered <- pep_dat %>% filter(pep_dat$new<0.01)

#filtered2<-pep_dat %>% filter(pep_dat$pair %in% filtered$pair)

#plot2+geom_abline(slope = 1,intercept = 0)+geom_line(data=filtered2,aes(group=pair))

#plot2+geom_abline(slope = 1,intercept = 0)+geom_line(data=filtered2,aes(group=pair))+xlim(0,0.05)

dat_p3 <- data.frame(RT1=y[,1],RT2=y[,2], new1=new_pep[1:n], new2=new_pep[(n+1):(2*n)],old1=old_pep[1:n], old2=old_pep[(n+1):(2*n)],id=(z==matrix(rep(1:n,p),nrow = n)),ref_RT1=x1[,1],ref_RT2=x1[,2])
ggplot(data = dat_p3)+geom_point(aes(x=RT1,y=RT2, colour=id.1))

ggplot(data = dat_p3)+geom_point(aes(x=RT1,y=RT2, colour=new1<0.01))


ggplot(data = dat_p3)+geom_point(aes(x=RT1,y=RT2, colour=id.2))

ggplot(data = dat_p3)+geom_point(aes(x=RT1,y=RT2, colour=new2<0.01))

ggplot(data = dat_p3)+geom_point(aes(x=RT1,y=ref_RT1, colour=new1<0.01))
ggplot(data = dat_p3)+geom_point(aes(x=RT2,y=ref_RT2, colour=new1<0.01))
ggplot(data = dat_p3)+geom_point(aes(x=RT1,y=ref_RT1, colour=new1<old1))

```

# Tables

## Table 1

For table 1, we see that there is 1 falsely identified peptide with pep<0.01. And the model fails identify this one.


```{r create tables}
# Create Tables

#1. how many mislabels
#2. how many PEP > 0.01


#Table 1 

count <- matrix(0,nrow = 2, ncol = 2)
z_logis <- matrix(FALSE,nrow = n,ncol = p)

for (i in 1:nrow(y)) {
  for(j in 1:ncol(y)){
    count[1,1]=count[1,1]+as.numeric((i==z[i,j])&(pep_sim[i,j]>0.01))
    count[1,2]=count[1,2]+as.numeric((i!=z[i,j])&(pep_sim[i,j]>0.01))
    count[2,1]=count[2,1]+as.numeric((i==z[i,j])&(pep_sim[i,j]<0.01))
    count[2,2]=count[2,2]+as.numeric((i!=z[i,j])&(pep_sim[i,j]<0.01))
    if(i!=z[i,j]){
      z_logis[i,j]=TRUE
    } 
  }
}
table1 <- as.data.frame(count)

rownames(table1) <- c("PEP>0.01","PEP<0.01")
colnames(table1) <- c("True Match","False Match")
kable(table1, "html", caption = "Table before update")

id_points_1<-which((z_logis)&(pep_sim<0.01),arr.ind = TRUE)
id_points_1
z[id_points_1]
updated_id[id_points_1]

sum(new_pep>0.01 & old_pep<0.01)
old_pep[new_pep>0.01 & old_pep<0.01]
new_pep[new_pep>0.01 & old_pep<0.01]
id.points_2<- which(mtx_updated_pep>0.01 & pep_sim<0.01,arr.ind = TRUE)
id.points_2

which(mtx_updated_true_pep>0.01 & pep_sim<0.01,arr.ind = TRUE)



dat_p5 <- data.frame(RT1=y[,1],RT2=y[,id_points_1[1,2]], new1=mtx_updated_pep[,1], new2=mtx_updated_pep[,id_points_1[1,2]], ref_RT1=x1[,2],ref_RT2=x1[,id_points_1[1,2]])

ggplot(data = dat_p5,aes(x=RT1,y=RT2))+geom_point()+geom_point(data = dat_p5[id_points_1[1,1],], aes(x=RT1,y=RT2), colour = "red")+geom_path(data = data.frame(x=y1[,1],y=y1[,id_points_1[1,2]]),aes(x=x,y=y))
```

## Table 2

For table 2 there are two wrongly identified points. These are different from the ones in table 2


```{r table 2}
#Table 2 updated version of table 1
count <- matrix(0,nrow = 2, ncol = 2)

for (i in 1:nrow(y)) {
  for(j in 1:ncol(y)){
    count[1,1]=count[1,1]+as.numeric((updated_id[i,j]==z[i,j])&(mtx_updated_pep[i,j]>0.01))
    count[1,2]=count[1,2]+as.numeric((updated_id[i,j]!=z[i,j])&(mtx_updated_pep[i,j]>0.01))
    count[2,1]=count[2,1]+as.numeric((updated_id[i,j]==z[i,j])&(mtx_updated_pep[i,j]<0.01))
    count[2,2]=count[2,2]+as.numeric((updated_id[i,j]!=z[i,j])&(mtx_updated_pep[i,j]<0.01))
  }
}
table2 <- as.data.frame(count)
rownames(table2) <- c("PEP>0.01","PEP<0.01")
colnames(table2) <- c("True Match","False Match")

kable(table2, "html", caption = "Table after update")

wrongly_updated_id <- which((updated_id!=z)&(mtx_updated_pep<0.01), arr.ind = TRUE)
wrongly_updated_id
pep_sim[wrongly_updated_id]#pep larger than
updated_id[wrongly_updated_id]
mtx_updated_pep[wrongly_updated_id]
z[wrongly_updated_id]


dat_p4 <- data.frame(RT1=y[,2],RT2=y[,wrongly_updated_id[1,2]], new1=mtx_updated_pep[,2], new2=mtx_updated_pep[,wrongly_updated_id[1,2]], ref_RT1=x1[,2],ref_RT2=x1[,wrongly_updated_id[1,2]])

ggplot(data = dat_p4,aes(x=RT1,y=RT2))+geom_point()+geom_point(data = dat_p4[wrongly_updated_id[1,1],], aes(x=RT1,y=RT2), colour = "red")+geom_path(data = data.frame(x=y1[,2],y=y1[,wrongly_updated_id[1,2]]),aes(x=x,y=y))

```



Improvement:

1. Allow replicated id?

2. Better explanation of wrongly identified points

3. rownumber and peptideid relationship delve more





































